Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-05-04 Manuscript/manuscript.Rmd
manuscript edits

An ensemble of models was established for the purposes of systematically comparing the performances of different parameters in the framework (Figure 2). Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type for the feature space.

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the minority and majority sets respectively. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For the task of performance evaluation and prediction, the test set remained constant for all ensemble systems to account for an unbiased comparison.

@@

We compared the performances of random forest classification models implemented on TAD boundaries called using the arrowhead algorithm [@durand2016juicebox] from Hi-C data for 2 cell lines $C = \{\textit{GM12878}, \textit{K562}\}$, at four different resolutions $R = \{\textit{10 kb}, \textit{25 kb}, \textit{50 kb}, \textit{100 kb\}}$. [??? Bad spacing, for the whole paragraph. Why you are not using plain and simple Markdown?] For each classifier, we considered 3 separate predictor types as the feature space including $P = \{\textit{OC}, \textit{OP}, \textit{Distance}\}$. To assess the severity of class imbalance 4 re-sampling techniques were implemented prior to modelling, including $B = \{\textit{None}, \textit{RUS}, \textit{ROS}, \textit{SMOTE}\}$. A system produced from the ensemble framework was defined as: $S = \{\textit{c}, \textit{r}, \textit{p}, \textit{b}\}$, where $c \in C$, $r \in R$, $p \in P$, and $b \in B$. Letting $|X|$ denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $96 (= 2 \times 4 \times 3 \times 4)$ separate models.

>>>>

An ensemble framework was established for the purposes of systematically creating a variety of models that each included different sets of parameters described above (Figure 2). Each system in the framework is assumed to be composed of a model constructed from a particular data resolution, with a particular predictor type for the feature space. The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the minority and majority sets respectively. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For the task of performance evaluation and prediction, the test set remained constant for all ensemble systems to account for an unbiased comparison. Furthermore, this process was repeated, and performances were aggregated across each autosomal chromosome. Model results for Chromosome 9 was omitted from the analysis, for both cell lines, due to the sparcity of the contact matrix at the 10 kb resolution for the K562 cell line.

@@

We compared the performances of random forest classification models implemented on TAD boundaries called using the arrowhead algorithm [@durand2016juicebox] from Hi-C data for 2 cell lines $L = \{\textit{GM12878}, \textit{K562}\}$, at four different resolutions $R = \{\textit{10 kb}, \textit{25 kb}, \textit{50 kb}, \textit{100 kb\}}$. [??? Bad spacing, for the whole paragraph. Why you are not using plain and simple Markdown?] For each classifier, we considered 3 separate predictor types as the feature space including $P = \{\textit{OC}, \textit{OP}, \textit{Distance}\}$. To assess the severity of class imbalance 4 re-sampling techniques were implemented prior to modelling, including $B = \{\textit{None}, \textit{RUS}, \textit{ROS}, \textit{SMOTE}\}$. Each combination of the above parameters was repeated 21 times (number of autosomal chromosomes, excluding chromosome 9), $C = \{ 1, 2, ..., 22 \}$. A system produced from the ensemble framework was defined as: $S = \{\textit{c}, \textit{r}, \textit{p}, \textit{b}\}$, where $l \in L$, $r \in R$, $p \in P$, $b \in B$, and $c \in C$. Letting $|X|$ denote the cardinality of each set, we evaluated $|L| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $2016 (= 2 \times 4 \times 3 \times 4 \times 21)$ separate models.
